---
title: "GWAS_apply"
output: html_document
---

```{r}
if (!file.exists("GAPIT_Tutorial_Data.zip"))
{
  download.file("http://zzlab.net/GAPIT/GAPIT_Tutorial_Data.zip", destfile = "GAPIT_Tutorial_Data.zip")
  unzip("GAPIT_Tutorial_Data.zip")
}
download.file("http://zzlab.net/GAPIT/data/CROP545_Covariates.txt", destfile = "CROPS545_Covariates.txt")
download.file("http://zzlab.net/GAPIT/data/CROP545_Phenotype.txt", destfile = "CROPS545_Phenotype.txt")
# Import the GAPIT demo data genotypes
gt_scan <- data.frame(read.table("GAPIT_Tutorial_Data/mdp_numeric.txt", header = T, stringsAsFactors = F, sep = "\t", nrows = 1))
classes <- sapply(gt_scan, class)
genotypes <- data.frame(read.table("GAPIT_Tutorial_Data/mdp_numeric.txt", header = T, row.names = 1, colClasses = classes, stringsAsFactors = F, sep = "\t"))

GM <- read.table("GAPIT_Tutorial_Data/mdp_SNP_information.txt", header = T, stringsAsFactors = F, sep = "\t")
CV <- read.table("CROPS545_Covariates.txt", header = T, stringsAsFactors = F, sep = "\t")
phenotypes <- read.table("CROPS545_Phenotype.txt", header = T, stringsAsFactors = F, sep = "\t")


#Load dependencies
if (!require("pacman")) install.packages("pacman")
pacman::p_load(mgvc,ggplot2,knitr,gridExtra,kableExtra,data.table)

source("R/Function_qqplot.R")
source("R/Function_manhattan_QTN.R")
source("R/Function_manhattan.R")
source("R/FUNCTION_powerFDRtype1error.R")
```

# Question 3

## Demo with QTN
```{r, fig.show='hide'}
#Create QTNs
n=nrow(genotypes)
m=ncol(genotypes)
NQTN=10
#Show the positions of the QTN
QTN.position=sample(m,NQTN,replace=F)
phenotypes_n_1=phenotypes[,2]
test=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,QTN.position,plots=TRUE,messages=TRUE)
hist(test$power.fdr.type1error.res$power)
```


## Demo without QTN
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(y=phenotypes_n_1,X=genotypes,C=CV,GM=GM,PCA.M = 3,plots=TRUE,messages=TRUE)
```

## Demo no plot
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,plots=FALSE,messages=TRUE)
```

## Demo no plot nor CV
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(y=phenotypes_n_1,X=genotypes,GM=GM,PCA.M = 3,plot=FALSE,messages=TRUE)
```

## Demo no plot nor CV
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(y=phenotypes_n_1,X=genotypes,GM=GM,PCA.M = 3,plot=FALSE,messages=TRUE)
```

## Demo with all parameters, plots and messages and returns a csv, specified trait name
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,QTN.position,plot=TRUE,messages=TRUE,print=TRUE,trait="height")
```

## Demo with all parameters, plots and messages and returns a csv, with a specified cutoff
```{r, fig.show='hide'}
phenotypes_n_1=phenotypes[,2]
test=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,QTN.position,plot=TRUE,messages=TRUE,print=TRUE,cutoff = 0.05)
test$power.fdr.type1error.res$FP.fdr.power
test$power.fdr.type1error.res$TP.fdr.power
```

# Question 4

## Simulated Data
```{r}
source("http://www.zzlab.net/StaGen/2020/R/G2P.R")
source("http://www.zzlab.net/StaGen/2020/R/GWASbyCor.R")
NQTN = 10 #specify number of QTN
h2 = 0.75 #heritability
alpha = 0.6 #alpha value
distribution = "normal" #specify distribution
set.seed(1337)
gcrep=replicate (30,{
  mySim=G2P(genotypes, h2, alpha, NQTN, distribution) #phenotype simulation
  p= as.vector(GWASbyCor(X=genotypes,y=phenotypes_n_1))#GWAS by using the correlation methods
  pres=power.fdr(p, QTN.position=QTN.position,cutoff=NULL)
  manhattan_plot(GM,p,cutoff=NULL,QTN.position,FP=pres$FP.fdr.power,TP=pres$TP.fdr.power, trait = "unknown",messages=TRUE)
  qq_plot(GM,p,QTN.position, trait = "unknown")
  })
means=mean(gcrep)
stdevs=sd(gcrep)
out4 <- rbind(means,stdevs)
out4 <- as.data.frame(out4)
colnames(out4) <- c("QTNs in the Top 10")
rownames(out4)<-c("Averages","Standard Deviations")
kable(out4) 
```

```{r}
NQTN = 10 #specify number of QTN
h2 = 0.75 #heritability
alpha = 0.6 #alpha value
distribution = "normal" #specify distribution
set.seed(1337)
n=nrow(genotypes)
m=ncol(genotypes)
NQTN=10
#Show the positions of the QTN
QTN.position=sample(m,NQTN,replace=F)

garep=replicate(30,{
  mySim=G2P(genotypes, h2, alpha, NQTN, distribution)
  function(y=NULL, X=NULL, C=NULL, GM=NULL, PCA.M=3, QTN.position=NULL, cutoff=NULL,plots=FALSE,messages=FALSE,print=FALSE)
  test_rep=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,QTN.position=QTN.position, plot=TRUE)
  N_detected=length(test_rep$True.Positive)
  N_output=c()
  N_output=c(N_detected)
  })
means=mean(garep)
stdevs=sd(garep)
out5 <- rbind(means,stdevs)
out5 <- as.data.frame(out5)
colnames(out5) <- c("True Positives")
rownames(out5)<-c("Averages","Standard Deviations")
kable(out5) 

T10r=index[1:10]
  detected=intersect(T10r,mySim$QTN.position)#find the position of top10 detected QTN
  N_detected= length(detected) # the number of QTNs among top ten associated SNPs
  N_output=c()
  N_output=c(N_detected)#combine the output into matrix
mean(test_rep$power.fdr.type1error.res$power)  
mean(test_rep$power.fdr.type1error.res$fdr) 
mean(test_rep$power.fdr.type1error.res$type1error) 
str(test_rep$P.value.res)
test_c=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3,QTN.position=QTN.position, plot=FALSE,cutoff = 0.05)
test_c$True.Positive
test_c$False.Positive
test_rep$False.Positive
pdr=power.fdr(test_rep$order.SNP.res,QTN.position)

fal.neg(test_rep$P.value.res,QTN.position,test_rep$cutoff.final.res)
```

# This code can be used for The bonus question

## Time Difference between GWAS with apply vs loop
```{r}
#Time BLINK
loop_start <- proc.time()
test=GWAStest(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3)
loop_end <- proc.time()
loop_elapsed <- loop_end[3] - loop_start[3]

#Time our function
apply_start <- proc.time()
test=GWASapply(phenotypes_n_1,genotypes,CV,GM,PCA.M = 3)
apply_end <- proc.time()
apply_elapsed <- apply_end[3] - apply_start[3]

# Consolidate times into a neat table
time_table <- data.frame(c("loop", "apply"), c(loop_elapsed, apply_elapsed))
names(time_table) <- c("Method", "Time(s)")
kable(time_table) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

# Calculate the percent improvement
perc_diff <- round((loop_elapsed - apply_elapsed) / loop_elapsed * 100,2)

print(paste("Apply is ", perc_diff , "% faster!"))
```
